Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-11-11 paper/main.tex
some edits

\subsubsection{CEGO}
%
We use the implementation of GECO provided by the
authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
Although it is never stated in the original paper, the implementation of CEGO
generates a set of initial solutions of size $\minit=10$ by means of a
max-min-distance sequential design: new solutions are added to the set
sequentially by maximizing the minimum distance to solutions already in the
set. These initial solutions are then evaluated on the true fitness function
and the result is used to build the initial surrogate
model. % \citet{ZaeStoBar2014:ppsn} uses \minit=10
Following the authors of
CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
the surrogate models with population size of 20, crossover rate of 0.5,
mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
interchange mutation (i.e., exchanging two randomly selected elements) and
cycle crossover for permutations. To generate each new solution, CEGO runs the
GA using the model as a surrogate of the true fitness function.  The authors of
CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
GA. However, with such a setting, a single run of CEGO with a budget of
$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of
computation time, which was infeasible for our study. Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 3 days for each run of CEGO. Hence, we set the budget of the
GA to $10^3$. Although a higher GA budget could improve the results of CEGO
reported here, our results are similar (and sometimes better) than the ones
reported in the original
papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to running CEGO
with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the increase
in runtime caused by doubling $\FEmax$ is much smaller than by increasing the
GA budget by a factor of $10$. Nevertheless, we caution that choosing a
different trade-off may lead to different conclusions and more work is needed
to understand CEGO parameters.

@@

Table~\ref{tab:results} shows results over all instances evaluated (excluding
the synthetic LOP instances). In particular, we show the mean fitness (and
standard deviation) of the best solution found at the end of each run. We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 72 hours, also the number of fitness evaluations reached by CEGO.

In the LOP, differences are mostly in favour of UMM, due to the time limit of
72 hours, which resulted in CEGO typically achieving less than 200
evaluations. In contrast, UMM was always able to reach 400 evaluations in less
than 16 hours. If UMM was allowed to reach the 72 hours limit, the differences
would be even stronger.

@@

The overall conclusion from Table~\ref{tab:results} is that there is not a
significant difference in performance between UMM and CEGO. For the larger LOP
instances, CEGO reaches the time limit of 72 hours before consuming the budget
of 400 evaluations and thus, obtains quite poor solutions compared to UMM,
which is at least an order of magnitude faster.

@@

 \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance. The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO. Mean runtime is measured in hours. If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier. This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}

>>>>

\subsubsection{CEGO}
%
We use the implementation of GECO provided by the
authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
Although it is never stated in the original paper, the implementation of CEGO
generates a set of initial solutions of size $\minit=10$ by means of a
max-min-distance sequential design: new solutions are added to the set
sequentially by maximizing the minimum distance to solutions already in the
set. These initial solutions are then evaluated on the true fitness function
and the result is used to build the initial surrogate
model. % \citet{ZaeStoBar2014:ppsn} uses \minit=10
Following the authors of
CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
the surrogate models with population size of 20, crossover rate of 0.5,
mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
interchange mutation (i.e., exchanging two randomly selected elements) and
cycle crossover for permutations. To generate each new solution, CEGO runs the
GA using the model as a surrogate of the true fitness function.  The authors of
CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
GA. However, with such a setting, a single run of CEGO with a budget of
$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of
computation time, which was infeasible for our study. Due to extremely long
runtimes of CEGO and constraints in our computing system, we set the budget of
the GA to $10^3$ and we stopped each run of CEGO after 144 hours of CPU-time.
Although a higher GA budget could improve the results of CEGO reported here,
our results are similar (and sometimes better) than the ones reported in the
original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that
doubling $\FEmax$ produces a smaller increase in runtime than multiplying the
GA budget by a factor of $10$. Nevertheless, we caution that choosing a
different trade-off may lead to different conclusions and more work is needed
to understand CEGO parameters.

@@

Table~\ref{tab:results} shows results over all instances evaluated (excluding
the synthetic LOP instances). In particular, we show the mean fitness (and
standard deviation) of the best solution found at the end of each run. We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 144 hours, also the number of fitness evaluations reached by CEGO.

In the LOP, differences are mostly in favour of UMM, due to the time limit of
144 hours, which resulted in CEGO typically achieving less than 200
evaluations. In contrast, UMM was always able to reach 400 evaluations in less
than 16 hours. If UMM was allowed to reach the 144 hours limit, the differences
would be even stronger.

@@

The overall conclusion from Table~\ref{tab:results} is that there is not a
significant difference in performance between UMM and CEGO. For the larger LOP
instances, CEGO reaches the time limit of 144 hours before consuming the budget
of 400 evaluations and thus, obtains quite poor solutions compared to UMM,
which is at least an order of magnitude faster.

@@

 \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance. The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO. Mean runtime is measured in hours. If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 144 hours, it was terminated earlier. This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}
